---
title: "Tensorize"
author: "Patrick Brophy"
date: "6/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(fst)
library(ggplot2)
source("PeptideSequence2Tensor.R")
source("PeptideFragments2Tensor.R")
```

#Import Data
This markdown picks up where "Prepare_NIST_msp.Rmd" concludes. The first step is to import the serialized .fst data generated at the end of "Prepare_NIST_msp.Rmd". Then, we convert it to a data.table.

```{r}
data <- read_fst(path = "human_consensus_final_true_lib.fst")
data <- as.data.table(data)
```

#Precursor Distributions: Charges and Sequence Lengths 
```{r eval=FALSE, include=FALSE}
scientific <- function(x){
    ifelse(x==0, "0", parse(text=gsub("[+]", "", gsub("e", " %*% 10^", scales::scientific_format()(x)))))
}

ggplot(data = data, aes(x = seqLen))+
  geom_histogram(binwidth = 1)+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        axis.text = element_text(color = "black"))+
  scale_y_continuous(expand = c(0,0), label = scientific)+
  labs(title = "Distribution of Peptides by Sequence Length", x = "Sequence Length")

ggplot(data = data, aes(x = Charge))+
  geom_histogram(binwidth = 1)+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        axis.text = element_text(color = "black"))+
  scale_y_continuous(expand = c(0,0), label = scientific)+
  labs(title = "Distribution of Peptides by Sequence Length", x = "Precursor Charge State")

ggplot(data = data, aes(x = seqLen))+
  geom_histogram(binwidth = 1)+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        axis.text = element_text(color = "black"))+
  scale_y_continuous(expand = c(0,0), label = scientific)+
  labs(title = "Distribution of Peptides by Sequence Length", 
       subtitle = "Subset by Charge State", 
       x = "Sequence Length")+
  facet_wrap("Charge", scales = "free")
```

#Subset Data
The approach for the initial test is to train a model to predict fragment ion intensity from integer encoded peptide sequences of a single charge state. The fragment ions are then either "y" or "b" ions with a charge less than or equal to that of the input peptide sequences.

##Precursor ion subsetting
```{r}
#Limit Precursor to charge == 2 & length <= 30
maxSequenceLength <- 30
chargeState <- 2

data <- data[seqLen <= maxSequenceLength, ]
data <- data[Charge == chargeState, ]

#Index by name and charge to provide unique identifier for precursor species 
#that could have multiple charge states but the same name (sequence). Not the case here... 
data[, name_charge_index := as.numeric(as.factor(paste0(Name, Charge)))]

numUniquePrecursors <- length(unique(data$name_charge_index))
```

##Fragment Ion 
```{r}
#Index fragment ions by unique y/b charge state combinations 
data[, fragType_chargeCode := paste0(ionType, fragChargeState)]
fragType_chargeCode_lookup <- sort(unique(data$fragType_chargeCode))
data[, fragType_chargeCode := match(x = fragType_chargeCode, table = fragType_chargeCode_lookup)]

#Number of fragments per unique precursor
data[, nFragments := max(.N), by = name_charge_index]

#Normalize ion intensity
data[, intensity_norm := intensity/max(intensity)]
```

#Tensorize 
The model will take as an input an array with dimensions (`r numUniquePrecursors`, `r maxSequenceLength`, 1). Each row is a unique precursor ion and each column is a location within the sequence. The array is filled with integers corresponding to the amino acid single letter codes.

The model will be trained on output data consisting of arrays with dimensions (`r numUniquePrecursors`, `r maxLength`, `r length(fragType_chargeCode_lookup)`). For a single peptide with a given charge state, this corresponds to a matrix  where the number of rows equals to the length of the longest peptide in the set; the number of columns equals the number of various ion types allowed by prior filtering: `r fragType_chargeCode_lookup`.

##Input Data Conversion - sequences at a given charge state
```{r}
#Convert all aa sequences into matrix grids
aa <- c("G", "A", "L", "M", "F", "W", "K", "Q", "E", "S", "P", "V", "I", "C", "Y", "H", "R", "N", "D", "T")
aa <- sort(aa)

#Subset the data.table by selecting unique sequence/charge combinations
data_unique <- unique(data[,list(name_charge_index, Name, Charge)])

input_integerSequences <- sequence2tensor_intSequence(uniqueDt = data_unique, aa = aa)

image(t(input_integerSequences[1:10,,1]), 
      y = c(1:10), 
      x = c(1:maxLength), 
      col = rainbow(n = 21), 
      xlab = "Sequence Index", 
      ylab = "Peptide Index")
```

##Output Data Conversion - product ions from a sequence at a given charge state
```{r}
output_fragmentEncodings <- fragments2tensor(fragmentDt = data, uniqueDt = data_unique)
image((output_fragmentEncodings[1,,]), 
      y = c(1:4), 
      x = c(1:maxLength), 
      col = hcl.colors(n = 100), 
      xlab = "Sequence Index", 
      ylab = "Fragment Type")

decode<- decodeFragArray(fragmentArray = output_fragmentEncodings[1:10,,])

#Ensure decodings match
decode_pep <- decode[sequence == "AATALLEAGLAR" & intensity > 0, .(fragment_type, intensity, fragment_charge)]
decode_pep <- decode_pep[order(decode_pep$fragment_type)]

data_pep <- data[Name == "AATALLEAGLAR" & intensity_norm > 0, .(goodAnnotation , intensity_norm, fragChargeState)]
data_pep <- data_pep[order(data_pep$goodAnnotation)]
```

```{r}
save(list = c("input_integerSequences", "output_fragmentEncodings"), file = "Tensors.Rdat")

```










